---
title: "2 comparing climate data"
author: "Matthew Ross"
date: "2024-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(RcppRoll)

gldas_dir = 'data/GLDAS/'

if(!file.exists(gldas_dir)){
  dir.create('data')
  dir.create(gldas_dir)
}
```


# Assignment


For this assignment we are going to compare climate data from a single point 
versus a watershed averaged climate data. We will be working over the Yampa river
watershed, one of the last undammed watersheds in the USA. 

## Point-climate acquisition. 

Using the 1_climate_data_yojoa.Rmd as an example download at least two types of climate (wind/temp/etc...) GLDAS data for the
Yampa River above Elkhead Creek. This is the site of a USGS gage that has data
dating back to 2004. 

```{r}
site_info <- tibble(site_no = '09244490',
                    lat = 40.5180278,
                    long = -107.3997838,
                    name = 'Yampa_hayden')
```



## Watershed averaged climate data

Using climate engine, download the two same parameters but for the watershed.
The watershed is stored as a .geojson in the yampa folder. Note, you likely need
to convert the watershed to a ".shp" file before getting the data from climate 
engine. 


## Compare your climate data anyway you want

Make at least two plots comparing your point versus watershed-averaged climate data. 



